---
title: "spotify_clustering"
output: spotify_clustering
---
GUIDE FROM - https://www.kaggle.com/prathitshivade/spotify-timeseries-clustering
https://developer.spotify.com/documentation/web-api/reference/tracks/get-audio-features/
https://www.kaggle.com/yamaerenay/spotify-dataset-19212020-160k-tracks
```{r}
library(caret)
library(ggplot2)
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(gridExtra)
```

```{r}
spotify_df <- read.csv("./spotify_files/data.csv")
```

```{r}
summary(spotify_df)

```


Remember that key is a categorical variable
```{r}
names(spotify_df)
```
1 valence - positivity rating
2 year
3 acousticness - 
4 artists
5 danceability
6 duration_ms
7 energy
8 explicit
9 id
10 instrumentalness
11 key
12 liveness
13 loudness
14 mode
15 name
16 popularity
17 release_date
18 speechiness
19 tempo

##Test Run with all notable traits
will probably split this one up into groups of 3 so we can visualize everything
```{r}
spotify_df <- spotify_df[,c(-2,-4,-8,-9,-11,-14,-15,-17)]
#spotify_df$mode <- factor(spotify_df$mode)
spotify_norm_values <- preProcess(spotify_df, method = c("center", "scale"))
spotify_df_norm <- predict(spotify_norm_values, spotify_df)
```
```{r}
head(spotify_df_norm)
```



```{r}
km_2 <- kmeans(spotify_df_norm,2, nstart=10)
km_3 <- kmeans(spotify_df_norm,3, nstart=10)
km_4 <- kmeans(spotify_df_norm,4, nstart=10)
km_5 <- kmeans(spotify_df_norm,5, nstart=10)
```

```{r}
km_2
km_3
km_4
km_5
```

```{r}
set.seed(123)

# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(spotify_df_norm, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```


```{r}
# plots to compare
km_3 <- kmeans(spotify_df_norm,3, nstart=10)
plot_1 <- fviz_cluster(km_3, geom = "point", data = spotify_df_norm) + ggtitle("k = 3")

```

## This plot automatically performs PCA (principal component analysis) which reduces all the dimensions to the first 2 principal components that explain the most variance
```{r}
plot_1

```
```{r}
km_3
```
Category characteristics
1- high in danceability, high energy, explicit, lyrical music
2- positive, energetic, upbeat
3- Acoustic, low energy, instrumental

